Notes on Perturbational Quantities

### 2008-07-29 docs/pert_notes/paper.tex
made doc vi friendly

\frac{\partial\rho_0}{\partial t} &=& -\nabla\cdot(\rho_0 w_0 \eb_r) - \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
&=& -w_0\frac{\partial\rho_0}{\partial r} \underbrace{- \rho_0\frac{\partial w_0}{\partial r}} - \nabla\cdot(\eta_\rho \eb_r). \label{rho0 equation}

@@

Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from (\ref{rho equation}) to give perturbational density equation:

@@

\frac{\partial\rho'}{\partial t} &=& -\nabla\cdot(\rho'\Ub) - \nabla\cdot(\rho_0\Ubt) + \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
&=& -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub - \nabla\cdot(\rho_0\Ubt)} + \nabla\cdot(\eta_\rho \eb_r).\label{rho' equation}

@@

We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.  The underbraced term in (\ref{rho' equation}) is computed in modify\_scal\_force.  The $\eta_\rho$ term is not included, but this is correct because the procedure to compute $\rho$ at edges described in Section \ref{Computing rho at edges} makes use of an edge-centered, time-centered $\rho_0^{n+\half}$ that does not include an $\eta_\rho$ contribution either.

@@

We first compute $\rho_0^{(2)}$ using advect\_base (but not correct\_base).  We compute $\rho$ on edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at edges. For the non-radial faces, we can directly add $\rho_0^{n+\half}$ since it is a cell-centered quantity.  For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ at radial faces.

@@

We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term is set to zero, and therefor we set is\_conservative = T to alter the discretization in make\_edge\_scal.

@@

To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.  However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was referred to in Section \ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed with a Godunov method in make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the underbraced term in (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ term is not included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.

@@

\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
&=& -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.\label{rhoh equation}

@@

\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T + \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi + (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k + \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.

@@

\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} + \overline{\rho H_{\rm ext}} \nonumber \\
&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} + \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} + \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}

@@

We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.  Then we will add the two answers together to get $\rho h$ at edges.  This is different from the density, in which we averaged an ``old'' and ``new'' base state density to get $\rho_0$ at edges.  By computing $\rho h$ at edges in this way, we avoid having to compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a correct\_base and diffuse\_base function for $\rho h$.\\

>>>>

\frac{\partial\rho_0}{\partial t} &=& -\nabla\cdot(\rho_0 w_0 \eb_r) - 
\nabla\cdot(\eta_\rho \eb_r) \nonumber \\
&=& -w_0\frac{\partial\rho_0}{\partial r} \underbrace{- 
\rho_0\frac{\partial w_0}{\partial r}} 
- \nabla\cdot(\eta_\rho \eb_r). \label{rho0 equation}

@@

Note that $\Ub = \Ubt + w_o\eb_r$.  Subtract (\ref{rho0 equation}) from 
(\ref{rho equation}) to give perturbational density equation:

@@

\frac{\partial\rho'}{\partial t} &=& -\nabla\cdot(\rho'\Ub) - \nabla\cdot(\rho_0\Ubt) 
+ \nabla\cdot(\eta_\rho \eb_r) \nonumber \\
&=& -\Ub\cdot\nabla\rho' \underbrace{- \rho'\nabla\cdot\Ub - \nabla\cdot(\rho_0\Ubt)} 
+ \nabla\cdot(\eta_\rho \eb_r).\label{rho' equation}

@@

We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.  The underbraced 
term in (\ref{rho' equation}) is computed in modify\_scal\_force.  The $\eta_\rho$ term 
is not included, but this is correct because the procedure to compute $\rho$ at edges 
described in Section \ref{Computing rho at edges} makes use of an edge-centered, 
time-centered $\rho_0^{n+\half}$ that does not include an $\eta_\rho$ contribution either.

@@

We first compute $\rho_0^{(2)}$ using advect\_base (but not correct\_base).  We compute 
$\rho$ on edges by adding $\rho_0^{n+\half} = (\rho_0^n + \rho_0^{(2)}) / 2$ to $\rho'$ at 
edges. For the non-radial faces, we can directly add $\rho_0^{n+\half}$ since it is a 
cell-centered quantity.  For the radial faces, we interpolate to obtain $\rho_0^{n+\half}$ 
at radial faces.

@@

We predict $\rho$ to edges using make\_edge\_scal in scalar\_advance.  The forcing term 
is set to zero, and therefore we set is\_conservative = T to alter the discretization 
in make\_edge\_scal.

@@

To compute $\eta_\rho$, we need $\rho_0$ at edge-centered, time-centered locations.  
However, we do NOT use the same edge-centered, time-centered $\rho_0^{n+\half}$ that was 
referred to in Section \ref{Computing rho at edges}.  Instead, we use a $\rho_0$ computed 
with a Godunov method in make\_edge\_state\_1d.  We call make\_edge\_state\_1d using the 
underbraced term in (\ref{rho0 equation}) as the forcing.  The $\eta_\rho$ term is not 
included.\MarginPar{I think there is a bug in the code and the $\eta_\rho$ term needs to 
be included here.}  This gives us the so-called $\rho_0^{n+\half,{\rm pred}}$.

@@

\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt} 
+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext} \nonumber \\
&=& -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub -\nabla\cdot(\rho h \Ub) 
+ \frac{Dp_0}{Dt} + \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} 
+ \rho H_{\rm ext}.\label{rhoh equation}

@@

\frac{\partial T}{\partial t} = -\Ub\cdot\nabla T
+ \frac{1}{\rho c_p}\left\{(1-\rho h_p)\left[\psi
+ (\Ubt\cdot\eb_r)\frac{\partial p_0}{\partial r}\right] - \sum_k\rho\xi_k\dot\omega_k 
+ \rho H_{\rm nuc} + \rho H_{\rm ext}\right\}.

@@

\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] 
+ \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
+ \overline{\rho H_{\rm ext}} \nonumber \\
&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
+ \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
+ \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}

@@

We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.  Then 
we will add the two answers together to get $\rho h$ at edges.  This is different from 
the density, in which we averaged an ``old'' and ``new'' base state density to get 
$\rho_0$ at edges.  By computing $\rho h$ at edges in this way, we avoid having to 
compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a 
correct\_base and diffuse\_base function for $\rho h$.\\
